---
title: "Submission"
---

# Data Import

```{r}
library(tidyverse)
library(lubridate)
library(readr)
library(ranger)
```

```{r}
challenge_set <- read_csv("challenge_set.csv")
```

```{r}
submission_set <- read_csv("final_submission_set.csv")
```

```{r}
challenge_set <- rbind(challenge_set, submission_set)
```

### Pull in Plane OEW/MTOW

```{r}
aircraft_oew_mtow <- read_csv("aircraft_oew_mtow.csv")
```

```{r}
challenge_set <- merge(x = challenge_set, y = aircraft_oew_mtow, by = "aircraft_type", all.x = TRUE)
```

### Quick Linear Regression for TOW by Flown Distance and Aircraft Type

```{r}
# Fit linear regression models by aircraft_type and get predicted values
challenge_set <- challenge_set %>%
  group_by(aircraft_type) %>%
  do({
    model <- lm(tow ~ flown_distance, data = na.omit(.))
    data.frame(., lm_predicted_tow = predict(model, newdata = .))
  }) %>%
  ungroup()

# Fit linear regression models by aircraft_type and extract R²
r_squared_values <- challenge_set %>%
  group_by(aircraft_type) %>%
  do(model = lm(tow ~ flown_distance, data = na.omit(.))) %>%
  summarise(aircraft_type,
            lm_predicted_tow_r_squared = summary(model)$r.squared
            )

challenge_set <- merge(x = challenge_set, y = r_squared_values, by = "aircraft_type", all.x = TRUE)
```

### Pull in Trajectories

```{r}
combined_trajectories <- read_csv("combined_trajectories_v5.csv")

# Drop extraneous columns
combined_trajectories <- combined_trajectories %>% select(-median_track, -.names, -count_trajectory_observations)
```

```{r}
# Clean Inf: impute with median
combined_trajectories <- combined_trajectories %>%
  mutate(across(where(is.numeric), ~ ifelse(is.infinite(.), median(., na.rm = TRUE), .)))

# Clean Negatives
neg_check_columns <- c("median_altitude", "p25_altitude", "p75_altitude", "max_altitude")

# Replace negative values with the column median
combined_trajectories <- combined_trajectories %>%
  mutate(across(all_of(neg_check_columns), ~ ifelse(. < 0, median(., na.rm = TRUE), .)))
```

```{r}
challenge_set <- merge(x = challenge_set, y = combined_trajectories, by = "flight_id", all.x = TRUE)
```

## Data Cleansing

### Clean up Null Trajectories

```{r}
# Impute NA to Zero
challenge_set$avg_headwind_tailwind <- ifelse(is.na(challenge_set$avg_headwind_tailwind), 0, challenge_set$avg_headwind_tailwind)
challenge_set$avg_humidity <- ifelse(is.na(challenge_set$avg_humidity), 0, challenge_set$avg_humidity)

# Impute NA to Median
# Impute NA to Median for all numeric columns except 'TOW'
challenge_set <- challenge_set %>%
  mutate(across(
    where(is.numeric) & !c('tow'),  # Select numeric columns excluding 'TOW'
    ~ ifelse(is.na(.), median(., na.rm = TRUE), .)
  ))
```

### Clean Up great_circle_distance

```{r}
challenge_set$great_circle_distance <- ifelse(challenge_set$great_circle_distance < 30, challenge_set$flown_distance, challenge_set$great_circle_distance)
```

# Modeling

## Feature Engineering

### Cluster Plane Models

```{r}
# aircraft_type_summary <- challenge_set %>% 
#   group_by(aircraft_type) %>%
#   summarize(
#     count = n(),
#     distinct_operators = n_distinct(airline),
#     median_flown_distance = median(flown_distance),
#     min_flown_distance = min(flown_distance),
#     max_flown_distance = max(flown_distance),
#     median_flown_velocity = median(flight_duration) / median(flown_distance),
#     max_flown_velocity = max(flight_duration / flown_distance),
#     median_tow = median(tow, na.rm = TRUE),
#     p25_tow = quantile(tow, 0.25, na.rm = TRUE)
#   )
# 
# # do kmeans
# set.seed(123) 
# kmeans_result <- aircraft_type_summary %>% 
#   select(count, distinct_operators, median_flown_distance, median_flown_velocity, median_tow) %>% 
#   scale() %>% 
#   kmeans(centers = 5)
# 
# # Add the cluster labels to the summarized data
# aircraft_type_summary$cluster = kmeans_result$cluster
# 
# # Write to file for use later
# aircraft_type_summary %>% 
#   select("aircraft_type", "cluster", "median_tow", "median_flown_distance", "count") %>% 
#   arrange(cluster) %>% 
#   write_csv("aircraft_type_clusters_kmeans.csv")
# 
# # Plot clusters
# ggplot(aircraft_type_summary, aes(x = median_flown_distance, y = median_tow, size = log(count), color = factor(cluster))) +
#   geom_point() +
#   theme_minimal() + 
#   labs(
#     title = "K-means Clustering of Aircraft Type", 
#     x = "Median Flown Distance", 
#     y = "Median TOW"
#   )
```

### Get Aircraft Type Clusters

```{r}
aircraft_type_clusters_kmeans <- read_csv("aircraft_type_clusters_kmeans.csv") %>%
  select("aircraft_type", "cluster") %>% 
  rename("aircraft_type_cluster" = "cluster")

aircraft_type_clusters_kmeans$aircraft_type_cluster = as.factor(aircraft_type_clusters_kmeans$aircraft_type_cluster)

challenge_set <- merge(x = challenge_set, y = aircraft_type_clusters_kmeans, by = "aircraft_type", all.x = TRUE)
```

### Get Airport-Centric Data

```{r}
airports <- read_csv("airports.csv")

airports <- airports %>% select("ident", "gps_code", "latitude_deg", "longitude_deg", "elevation_ft")
```

#### ADEP Airport Info

```{r}
# First join on "ident"
challenge_set <- merge(x = challenge_set, y = airports, by.x = "adep", by.y = "ident", all.x = TRUE) %>%
  rename("latitude_adep" = "latitude_deg", "longitude_adep" = "longitude_deg", "elevation_ft_adep" = "elevation_ft")

# Find rows where the first join did not succeed (NA values) and join on "gps_code"
challenge_set_join2 <- merge(x = challenge_set, y = airports, by.x = "adep", by.y = "gps_code", all.x = TRUE)

# Only keep rows where the first join resulted in NA values, but the second join worked
challenge_set <- challenge_set %>%
  mutate(
    latitude_adep = coalesce(latitude_adep, challenge_set_join2$latitude_deg),
    longitude_adep = coalesce(longitude_adep, challenge_set_join2$longitude_deg),
    elevation_ft_adep = coalesce(elevation_ft_adep, challenge_set_join2$elevation_ft, 0)
  )
```

#### ADES Airport Info

```{r}
# First join on "ident"
challenge_set <- merge(x = challenge_set, y = airports, by.x = "ades", by.y = "ident", all.x = TRUE) %>%
  rename("latitude_ades" = "latitude_deg", "longitude_ades" = "longitude_deg", "elevation_ft_ades" = "elevation_ft")

# Find rows where the first join did not succeed (NA values) and join on "gps_code"
challenge_set_join2 <- merge(x = challenge_set, y = airports, by.x = "ades", by.y = "gps_code", all.x = TRUE)

# Only keep rows where the first join resulted in NA values, but the second join worked
challenge_set <- challenge_set %>%
  mutate(
    latitude_ades = coalesce(latitude_ades, challenge_set_join2$latitude_deg),
    longitude_ades = coalesce(longitude_ades, challenge_set_join2$longitude_deg),
    elevation_ft_ades = coalesce(elevation_ft_ades, challenge_set_join2$elevation_ft, 0)
  )
```

```{r}
# Haversine formula function
haversine_distance <- function(lat1, lon1, lat2, lon2) {
  # Convert degrees to radians
  lat1 <- lat1 * pi / 180
  lon1 <- lon1 * pi / 180
  lat2 <- lat2 * pi / 180
  lon2 <- lon2 * pi / 180
  
  # Haversine formula
  dlat <- lat2 - lat1
  dlon <- lon2 - lon1
  a <- sin(dlat / 2)^2 + cos(lat1) * cos(lat2) * sin(dlon / 2)^2
  c <- 2 * atan2(sqrt(a), sqrt(1 - a))
  r <- 6371  # Radius of Earth in kilometers
  distance <- r * c
  
  return(distance)  # Return distance in KM
}
```

```{r}
challenge_set <- challenge_set %>% 
  mutate(
    great_circle_distance_adep_ades = coalesce(haversine_distance(latitude_ades, longitude_ades, latitude_adep, longitude_adep), great_circle_distance) * 0.539957,  # convert KM to NMI
    adep_ades_elevation_ft_diff = elevation_ft_adep - elevation_ft_ades
  )
```

### Collapse Rare Categories

#### Airline

```{r}
collapse_rare_categories_by_prop <- function(df, column, prop_threshold = 0.01, new_category = "Other") {
  # Calculate frequency of each category
  freq_table <- table(df[[column]])
  
  # Find categories that are below the proportion threshold
  rare_categories <- names(freq_table[freq_table / sum(freq_table) <= prop_threshold])
  
  # Replace rare categories with new category (e.g., "Other")
  df[[column]] <- ifelse(df[[column]] %in% rare_categories, new_category, df[[column]])
  
  return(df)
}

collapse_rare_categories_by_count <- function(df, column, min_count = 50, new_category = "Other") {
  # Calculate frequency of each category
  freq_table <- table(df[[column]])
  
  # Find categories that are below the absolute count threshold
  rare_categories <- names(freq_table[freq_table <= min_count])
  
  # Replace rare categories with new category (e.g., "Other")
  df[[column]] <- ifelse(df[[column]] %in% rare_categories, new_category, df[[column]])
  
  return(df)
}

# List of columns to collapse rare categories
columns_to_collapse_prop <- c("airline", "aircraft_type")

# Loop over each column and collapse rare categories
for (column in columns_to_collapse_prop) {
  challenge_set <- collapse_rare_categories_by_prop(challenge_set, column, prop_threshold = 0.0025, new_category = "Rare")
}

columns_to_collapse_abs <- c("adep", "ades", "country_code_ades", "country_code_adep")

# Loop over each column and collapse rare categories
for (column in columns_to_collapse_abs) {
  challenge_set <- collapse_rare_categories_by_count(challenge_set, column, min_count = 3, new_category = "Rare")
}

```

```{r}
# hist(challenge_set$flown_distance_over_great_circle,      
#      breaks = 30, 
#      col = "blue", 
#      border = "black")
```

```{r}
# challenge_set %>% 
#   arrange(desc(flown_distance_over_great_circle)) %>% 
#   head(10)
```

### Transform Features

```{r}
challenge_set <- challenge_set %>% 
  mutate(
    offblock_time_of_day = case_when(
      hour(actual_offblock_time) %in% c(6,7,8,9,10,11) ~ 'Morning',
      hour(actual_offblock_time) %in% c(12,13,14,15,16,17) ~ 'Afternoon',
      hour(actual_offblock_time) %in% c(18,19,20,21,22,23) ~ 'Evening',
      hour(actual_offblock_time) %in% c(0,1,2,3,4,5) ~ 'Night',
      .default = 'Unknown'
    ),
    offblock_is_peak = case_when(
      hour(actual_offblock_time) %in% c(7,8,17,18) ~ 'Peak',
      .default = 'Off Peak'
    ),
    arrival_time_of_day = case_when(
      hour(arrival_time) %in% c(6,7,8,9,10,11) ~ 'Morning',
      hour(arrival_time) %in% c(12,13,14,15,16,17) ~ 'Afternoon',
      hour(arrival_time) %in% c(18,19,20,21,22,23) ~ 'Evening',
      hour(arrival_time) %in% c(0,1,2,3,4,5) ~ 'Night',
      .default = 'Unknown'
    ),
    arrival_is_peak = case_when(
      hour(arrival_time) %in% c(7,8,17,18) ~ 'Peak',
      .default = 'Off Peak'
    ),
    day_of_week = wday(date, label = TRUE),
    is_weekend = if_else(wday(date, label = TRUE) %in% c('Sat', 'Sun'), TRUE, FALSE),
    week_of_year = as.factor(isoweek(date)),
    month_of_year = month(date, label = TRUE),
    is_domestic = if_else(country_code_adep == country_code_ades, TRUE, FALSE),
    flown_velocity = flown_distance / flight_duration,
    aircraft_family = substr(aircraft_type, 1, 3),
    distance_category = cut(flown_distance, breaks = c(0, 1000, 2700, Inf), labels = c("short", "medium", "long")),
    distance_bucket = coalesce(cut(flown_distance, 
                          breaks = seq(0, quantile(flown_distance, 0.999, na.rm = TRUE) + 500, by = 500), 
                          right = FALSE, 
                          include.lowest = TRUE), 'other'),
    flown_velocity_bucket = cut(flown_distance / flight_duration, breaks = seq(0, max(flown_distance / flight_duration, na.rm = TRUE) + 1, by = 0.5), right = FALSE, include.lowest = TRUE),
    relative_distance_adep_ades = cut(flown_distance / ave(flown_distance, adep, ades, FUN = mean), breaks = 10),
    relative_distance_aircraft_type = flown_distance / ave(flown_distance, aircraft_type, FUN = mean),
    relative_flown_velocity_aircraft_type = (flown_distance / flight_duration) / ave((flown_distance / flight_duration), aircraft_type, FUN = mean),
    relative_distance_adep_ades_aircraft_type = cut(flown_distance / ave(flown_distance, adep, ades, aircraft_type, FUN = mean), breaks = 10),
    relative_distance_airline = flown_distance / ave(flown_distance, airline, FUN = mean),
    relative_distance_aircraft_type_airline = flown_distance / ave(flown_distance, aircraft_type, airline, FUN = mean),
    cruise_duration = pmax(flight_duration - 30, 0),
    flight_minus_taxiout_duration_bucket = coalesce(cut(pmax(flight_duration - taxiout_time, 0), 
                                               breaks = seq(0, quantile(pmax(flight_duration - taxiout_time, 0), 0.999, na.rm = TRUE) + 30, by = 30),  # Adjust flight_minus_taxiout_duration_bucket based on 99.9th percentile
                                               right = FALSE, 
                                               include.lowest = TRUE), 'other'),
    departure_traffic = ave(flight_id, adep, date, FUN = length),
    arrival_traffic = ave(flight_id, ades, date, FUN = length),
    departure_traffic_squared = departure_traffic ^ 2,
    arrival_traffic_squared = arrival_traffic ^ 2,
    departure_traffic_bucket = cut(departure_traffic, breaks = 10),
    arrival_traffic_bucket = cut(arrival_traffic, breaks = 10),
    product_departure_traffic_arrival_traffic = departure_traffic * arrival_traffic,
    log_flight_duration = log1p(flight_duration),
    log_flown_distance = log1p(flown_distance),
    log_flown_distance_bucket = cut(log1p(flown_distance), breaks = 10),
    log_flown_velocity = log1p(flown_distance / flight_duration),
    flown_distance_squared = flown_distance ^ 2,
    flown_distance_sqrt = sqrt(flown_distance),
    flown_velocity_squared = (flown_distance / flight_duration) ^ 2,
    flown_velocity_sqrt = sqrt(flown_distance / flight_duration),
    taxiout_time_bucket = cut(taxiout_time, breaks = 10),
    taxiout_time_squared = taxiout_time ^ 2,
    taxiout_to_flight_duration_ratio = taxiout_time / flight_duration,
    taxiout_to_flight_duration_sqrt = sqrt(taxiout_time / flight_duration),
    log_great_circle_distance_adep_ades = log1p(great_circle_distance_adep_ades),
    great_circle_distance_squared = great_circle_distance_adep_ades ^ 2,
    great_circle_distance_sqrt = sqrt(great_circle_distance_adep_ades),
    log_median_altitude = log1p(median_altitude),
    median_altitude_squared = median_altitude ^ 2,
    log_median_groundspeed = log1p(median_groundspeed),
    altitude_range = p75_altitude - p25_altitude,
    groundspeed_range = p75_groundspeed - p25_groundspeed,
    altitude_range_bucket = cut(altitude_range, breaks = 10),
    groundspeed_range_bucket = cut(groundspeed_range, breaks = 10),
    flown_distance_over_great_circle = flown_distance / great_circle_distance_adep_ades,
    great_circle_distance_bucket = coalesce(cut(great_circle_distance_adep_ades, 
                                       breaks = seq(0, quantile(great_circle_distance_adep_ades, 0.999, na.rm = TRUE) + 500, by = 500),  # Adjust great_circle_distance_bucket based on 99.9th percentile
                                       right = FALSE, 
                                       include.lowest = TRUE), 'other'), 
    total_time_ascent_squared = total_time_ascent ^ 2,
    total_time_ascent_bucket = cut(total_time_ascent, breaks = 10),
    total_time_descent_squared = total_time_descent ^ 2,
    total_time_descent_bucket = cut(total_time_descent, breaks = 10),
    pct_total_time_ascent = total_time_ascent / flight_duration,
    pct_total_time_descent = total_time_descent / flight_duration,
    mtow_oew_diff = mtow - oew,
    pct_max_flown_distance_by_aircraft_type = cut((flown_distance / ave(flown_distance, aircraft_type, FUN = max)), breaks = 10),
    headwind_tailwind_bucket = cut(avg_headwind_tailwind, breaks = 10),
    avg_humidity_sqrt = sqrt(avg_humidity),
    avg_humidity_bucket = cut(avg_humidity_sqrt, breaks = 10),
    flown_distance_pct_range = flown_distance / range,
    predicted_tow_category = cut(lm_predicted_tow, breaks = quantile(lm_predicted_tow, probs = seq(0, 1, 0.1)), include.lowest = TRUE),
    lm_predicted_tow_squared = lm_predicted_tow^2,
    lm_predicted_tow_cubed = lm_predicted_tow^3,
    log_lm_predicted_tow = log(lm_predicted_tow + 1),
    r_squared_category = cut(lm_predicted_tow_r_squared,
                                         breaks = c(-Inf, 0, 0.5, 0.7, 0.9, 1),
                                         labels = c("Very Poor", "Poor", "Average", "Good", "Excellent"),
                                         include.lowest = TRUE)

  )
```

### Make Interaction Variables

```{r}
# aircraft type cluster
challenge_set$interaction_aircraft_type_cluster_log_great_circle_distance_adep_ades <- interaction(challenge_set$aircraft_type_cluster, challenge_set$log_great_circle_distance_adep_ades)
challenge_set$interaction_aircraft_type_cluster_log_flown_distance <- interaction(challenge_set$aircraft_type_cluster, challenge_set$log_flown_distance)
challenge_set$interaction_aircraft_type_cluster_log_flight_duration <- interaction(challenge_set$aircraft_type_cluster, challenge_set$log_flight_duration)
challenge_set$interaction_aircraft_type_cluster_flown_distance_squared <- interaction(challenge_set$aircraft_type_cluster, challenge_set$flown_distance_squared)
challenge_set$interaction_aircraft_type_cluster_cruise_duration <- interaction(challenge_set$aircraft_type_cluster, challenge_set$cruise_duration)
challenge_set$interaction_aircraft_type_cluster_total_time_ascent <- interaction(challenge_set$aircraft_type_cluster, challenge_set$total_time_ascent)
challenge_set$interaction_aircraft_type_cluster_airline <- interaction(challenge_set$airline, challenge_set$aircraft_type_cluster)
challenge_set$interaction_aircraft_type_cluster_adep <- interaction(challenge_set$adep, challenge_set$aircraft_type_cluster)
challenge_set$interaction_aircraft_type_cluster_ades <- interaction(challenge_set$ades, challenge_set$aircraft_type_cluster)
challenge_set$interaction_aircraft_type_cluster_distance_bucket <- interaction(challenge_set$distance_bucket, challenge_set$aircraft_type_cluster)
challenge_set$interaction_aircraft_type_cluster_flown_velocity_bucket <- interaction(challenge_set$aircraft_type_cluster, challenge_set$flown_velocity_bucket)
challenge_set$interaction_aircraft_type_cluster_median_altitude <- interaction(challenge_set$aircraft_type_cluster, challenge_set$median_altitude)
challenge_set$interaction_aircraft_type_cluster_log_median_altitude <- interaction(challenge_set$aircraft_type_cluster, challenge_set$log_median_altitude)
challenge_set$interaction_aircraft_type_cluster_p25_altitude <- interaction(challenge_set$aircraft_type_cluster, challenge_set$p25_altitude)
challenge_set$interaction_aircraft_type_cluster_max_altitude <- interaction(challenge_set$aircraft_type_cluster, challenge_set$max_altitude)
challenge_set$interaction_aircraft_type_cluster_total_time_ascent <- interaction(challenge_set$total_time_ascent, challenge_set$aircraft_type_cluster)
challenge_set$interaction_aircraft_type_cluster_total_time_descent <- interaction(challenge_set$total_time_descent, challenge_set$aircraft_type_cluster)
challenge_set$interaction_aircraft_type_cluster_total_time_climb_4 <- interaction(challenge_set$total_time_climb_4, challenge_set$aircraft_type_cluster)
challenge_set$interaction_aircraft_type_cluster_percentage_time_climb_4 <- interaction(challenge_set$percentage_time_climb_4, challenge_set$aircraft_type_cluster)
challenge_set$interaction_aircraft_type_cluster_month_of_year <- interaction(challenge_set$aircraft_type_cluster, challenge_set$month_of_year)
challenge_set$interaction_aircraft_type_cluster_country_code_ades <- interaction(challenge_set$aircraft_type_cluster, challenge_set$country_code_ades)
challenge_set$interaction_aircraft_type_cluster_log_flown_velocity <- interaction(challenge_set$aircraft_type_cluster, challenge_set$log_flown_velocity)
challenge_set$interaction_aircraft_family_distance_category <- interaction(challenge_set$aircraft_family, challenge_set$distance_category)
challenge_set$interaction_aircraft_type_cluster_product_departure_traffic_arrival_traffic <- interaction(challenge_set$aircraft_type_cluster, challenge_set$product_departure_traffic_arrival_traffic)
challenge_set$interaction_aircraft_type_distance_bucket_airline <- interaction(challenge_set$distance_bucket, challenge_set$aircraft_type, challenge_set$aircraft_type)

# aircraft type
challenge_set$interaction_aircraft_type_distance_bucket <- interaction(challenge_set$distance_bucket, challenge_set$aircraft_type)
challenge_set$interaction_aircraft_type_flight_minus_taxiout_duration_bucket <- interaction(challenge_set$flight_minus_taxiout_duration_bucket, challenge_set$aircraft_type)
challenge_set$interaction_aircraft_type_flown_velocity_bucket <- interaction(challenge_set$aircraft_type, challenge_set$flown_velocity_bucket)
challenge_set$interaction_aircraft_type_median_altitude <- interaction(challenge_set$aircraft_type, challenge_set$median_altitude)
challenge_set$interaction_aircraft_type_offblock_time_of_day <- interaction(challenge_set$offblock_time_of_day, challenge_set$aircraft_type)
challenge_set$interaction_aircraft_type_month_of_year <- interaction(challenge_set$aircraft_type, challenge_set$month_of_year)
challenge_set$interaction_aircraft_type_offblock_is_peak <- interaction(challenge_set$aircraft_type, challenge_set$offblock_is_peak)
challenge_set$interaction_aircraft_type_pct_max_distance_flown_by_aircraft_type <- interaction(challenge_set$aircraft_type, challenge_set$pct_max_flown_distance_by_aircraft_type)
challenge_set$interaction_aircraft_type_altitude_range_bucket <- interaction(challenge_set$aircraft_type, challenge_set$altitude_range_bucket)
challenge_set$interaction_aircraft_type_groundspeed_range_bucket <- interaction(challenge_set$aircraft_type, challenge_set$groundspeed_range_bucket)
challenge_set$interaction_aircraft_type_elevation_ft_adep <- interaction(challenge_set$aircraft_type, challenge_set$elevation_ft_adep)
challenge_set$interaction_aircraft_type_country_code_adep <- interaction(challenge_set$aircraft_type, challenge_set$country_code_adep)
challenge_set$interaction_aircraft_type_country_code_ades <- interaction(challenge_set$aircraft_type, challenge_set$country_code_ades)
challenge_set$interaction_aircraft_type_airline <- interaction(challenge_set$aircraft_type, challenge_set$airline)
challenge_set$interaction_aircraft_type_adep <- interaction(challenge_set$aircraft_type, challenge_set$adep)
challenge_set$interaction_aircraft_type_ades <- interaction(challenge_set$aircraft_type, challenge_set$ades)
challenge_set$interaction_aircraft_type_aircraft_family <- interaction(challenge_set$aircraft_type, challenge_set$aircraft_family)
challenge_set$interaction_aircraft_type_taxiout_time_bucket <- interaction(challenge_set$aircraft_type, challenge_set$taxiout_time_bucket)
challenge_set$interaction_aircraft_type_month_of_year <- interaction(challenge_set$aircraft_type, challenge_set$month_of_year)
challenge_set$interaction_aircraft_type_offblock_time_of_day <- interaction(challenge_set$aircraft_type, challenge_set$offblock_time_of_day)
challenge_set$interaction_aircraft_type_offblock_is_peak <- interaction(challenge_set$aircraft_type, challenge_set$offblock_is_peak)
challenge_set$interaction_aircraft_type_total_time_ascent_bucket <- interaction(challenge_set$aircraft_type, challenge_set$total_time_ascent_bucket)
challenge_set$interaction_aircraft_type_total_time_descent_bucket <- interaction(challenge_set$aircraft_type, challenge_set$total_time_descent_bucket)
challenge_set$interaction_aircraft_type_departure_traffic_bucket <- interaction(challenge_set$aircraft_type, challenge_set$departure_traffic_bucket)
challenge_set$interaction_aircraft_type_distance_category <- interaction(challenge_set$aircraft_type, challenge_set$distance_category)
challenge_set$interaction_aircraft_type_product_departure_traffic_arrival_traffic <- interaction(challenge_set$aircraft_type, challenge_set$product_departure_traffic_arrival_traffic)
challenge_set$interaction_aircraft_type_predicted_tow <- interaction(challenge_set$aircraft_type, challenge_set$lm_predicted_tow)


# airline
interaction_airline_flight_duration <- interaction(challenge_set$airline, challenge_set$flight_duration)
interaction_airline_great_circle_distance <- interaction(challenge_set$airline, challenge_set$great_circle_distance_adep_ades)
interaction_airline_flown_distance <- interaction(challenge_set$airline, challenge_set$flown_distance)
challenge_set$interaction_airline_aircraft_type <- interaction(challenge_set$airline, challenge_set$aircraft_type)
challenge_set$interaction_airline_distance_bucket <- interaction(challenge_set$distance_bucket, challenge_set$airline)
challenge_set$interaction_airline_distance_category <- interaction(challenge_set$distance_category, challenge_set$airline)
challenge_set$interaction_airline_flight_minus_taxiout_duration_bucket <- interaction(challenge_set$flight_minus_taxiout_duration_bucket, challenge_set$airline)
challenge_set$interaction_airline_offblock_time_of_day <- interaction(challenge_set$distance_bucket, challenge_set$offblock_time_of_day)
challenge_set$interaction_airline_product_departure_traffic_arrival_traffic <- interaction(challenge_set$airline, challenge_set$product_departure_traffic_arrival_traffic)
interaction_airline_offblock_is_peak <- interaction(challenge_set$airline, challenge_set$offblock_is_peak)
interaction_airline_month_of_year <- interaction(challenge_set$airline, challenge_set$month_of_year)
interaction_airline_day_of_week <- interaction(challenge_set$airline, challenge_set$day_of_week)
interaction_airline_distance_category <- interaction(challenge_set$airline, challenge_set$distance_category)
interaction_airline_country_code_adep <- interaction(challenge_set$airline, challenge_set$country_code_adep)
interaction_airline_country_code_ades <- interaction(challenge_set$airline, challenge_set$country_code_ades)
interaction_airline_departure_traffic <- interaction(challenge_set$airline, challenge_set$departure_traffic)
interaction_airline_taxiout_time <- interaction(challenge_set$airline, challenge_set$taxiout_time)
interaction_airline_total_time_ascent <- interaction(challenge_set$airline, challenge_set$total_time_ascent)
interaction_airline_flight_minus_taxiout_duration_bucket <- interaction(challenge_set$airline, challenge_set$flight_minus_taxiout_duration_bucket)
interaction_airline_flown_velocity_bucket <- interaction(challenge_set$airline, challenge_set$flown_velocity_bucket)
interaction_airline_altitude_range_bucket <- interaction(challenge_set$airline, challenge_set$altitude_range_bucket)
interaction_airline_avg_humidity_bucket <- interaction(challenge_set$airline, challenge_set$avg_humidity_bucket)
interaction_airline_distance_bucket <- interaction(challenge_set$airline, challenge_set$distance_bucket)
interaction_airline_headwind_tailwind_bucket <- interaction(challenge_set$airline, challenge_set$headwind_tailwind_bucket)


# adep 
challenge_set$interaction_adep_airline <- interaction(challenge_set$adep, challenge_set$airline)
challenge_set$interaction_adep_aircraft_type <- interaction(challenge_set$adep, challenge_set$aircraft_type)
challenge_set$interaction_adep_month_of_year <- interaction(challenge_set$adep, challenge_set$month_of_year)
challenge_set$interaction_adep_ades <- interaction(challenge_set$adep, challenge_set$ades)
challenge_set$interaction_adep_ades_airline <- interaction(challenge_set$interaction_adep_ades, challenge_set$airline)
challenge_set$interaction_adep_distance_bucket <- interaction(challenge_set$adep, challenge_set$distance_bucket)
challenge_set$interaction_adep_flown_distance_squared <- interaction(challenge_set$adep, challenge_set$flown_distance_squared)

# ades
challenge_set$interaction_ades_airline <- interaction(challenge_set$ades, challenge_set$airline)
challenge_set$interaction_ades_aircraft_type <- interaction(challenge_set$ades, challenge_set$aircraft_type)
challenge_set$interaction_ades_month_of_year <- interaction(challenge_set$ades, challenge_set$month_of_year)
challenge_set$interaction_country_code_ades_country_code_adep <- interaction(challenge_set$country_code_ades, challenge_set$country_code_adep)
challenge_set$interaction_country_code_ades_flown_distance_bucket <- interaction(challenge_set$country_code_ades, challenge_set$distance_bucket)
challenge_set$interaction_country_code_ades_flown_distance_squared <- interaction(challenge_set$country_code_ades, challenge_set$flown_distance_squared)

# time
challenge_set$interaction_day_of_week_departure_traffic <- interaction(challenge_set$day_of_week, challenge_set$departure_traffic)
challenge_set$interaction_month_of_year_aircraft_type_cluster_flown_distance_bucket <- interaction(challenge_set$month_of_year, challenge_set$aircraft_type_cluster, challenge_set$distance_bucket)
challenge_set$interaction_distance_bucket_offblock_time_of_day <- interaction(challenge_set$distance_bucket, challenge_set$offblock_time_of_day)

# great_circle_distance_bucket
challenge_set$interaction_great_circle_distance_bucket_aircraft_type_cluster <- interaction(challenge_set$great_circle_distance_bucket, challenge_set$aircraft_type_cluster)
challenge_set$interaction_great_circle_distance_bucket_aircraft_type <- interaction(challenge_set$great_circle_distance_bucket, challenge_set$aircraft_type)
challenge_set$interaction_great_circle_distance_bucket_flown_velocity_bucket <- interaction(challenge_set$great_circle_distance_bucket, challenge_set$flown_velocity_bucket)
challenge_set$interaction_great_circle_distance_bucket_country_code_ades <- interaction(challenge_set$great_circle_distance_bucket, challenge_set$country_code_ades)
challenge_set$interaction_great_circle_distance_bucket_median_altitude <- interaction(challenge_set$great_circle_distance_bucket, challenge_set$median_altitude)

# continuous variable interactions
challenge_set$interaction_log_great_circle_distance_adep_ades_log_median_altitude <- interaction(challenge_set$log_great_circle_distance_adep_ades, challenge_set$log_median_altitude)
challenge_set$interaction_distance_category_median_groundspeed <- interaction(challenge_set$distance_category, challenge_set$median_groundspeed)
challenge_set$interaction_total_time_ascent_aircraft_type <- interaction(challenge_set$total_time_ascent, challenge_set$aircraft_type)
challenge_set$interaction_total_time_descent_aircraft_type <- interaction(challenge_set$total_time_descent, challenge_set$aircraft_type)
#challenge_set$interaction_log_flown_velocity_log_median_altitude <- interaction(challenge_set$log_flown_velocity, challenge_set$log_median_altitude)
#challenge_set$interaction_log_flown_velocity_median_altitude_squared <- interaction(challenge_set$log_flown_velocity, challenge_set$median_altitude_squared)
challenge_set$interaction_altitude_range_bucket_distance_bucket <- interaction(challenge_set$altitude_range_bucket, challenge_set$distance_bucket)
challenge_set$interaction_groundspeed_range_distance_bucket <- interaction(challenge_set$groundspeed_range_bucket, challenge_set$distance_bucket)

# # mtow
# challenge_set$interaction_mtow_distance_bucket <- interaction(challenge_set$mtow, challenge_set$distance_bucket)
# challenge_set$interaction_mtow_great_circle_distance_bucket <- interaction(challenge_set$mtow, challenge_set$great_circle_distance_bucket)
# challenge_set$interaction_mtow_adep <- interaction(challenge_set$mtow, challenge_set$adep)
# challenge_set$interaction_mtow_ades <- interaction(challenge_set$mtow, challenge_set$ades)
# challenge_set$interaction_mtow_aircraft_type_cluster <- interaction(challenge_set$mtow, challenge_set$ades)
# challenge_set$interaction_mtow_cruise_duration <- interaction(challenge_set$mtow, challenge_set$cruise_duration)
# challenge_set$interaction_mtow_log_flown_distance <- interaction(challenge_set$mtow, challenge_set$log_flown_distance)
# challenge_set$interaction_mtow_log_flight_duration <- interaction(challenge_set$mtow, challenge_set$log_flight_duration)
# # oew
# challenge_set$interaction_oew_distance_bucket <- interaction(challenge_set$oew, challenge_set$distance_bucket)
# challenge_set$interaction_oew_great_circle_distance_bucket <- interaction(challenge_set$oew, challenge_set$great_circle_distance_bucket)
# challenge_set$interaction_oew_adep <- interaction(challenge_set$oew, challenge_set$adep)
# challenge_set$interaction_oew_ades <- interaction(challenge_set$oew, challenge_set$ades)
# challenge_set$interaction_oew_aircraft_type_cluster <- interaction(challenge_set$oew, challenge_set$ades)
# challenge_set$interaction_oew_cruise_duration <- interaction(challenge_set$oew, challenge_set$cruise_duration)
# challenge_set$interaction_oew_log_flown_distance <- interaction(challenge_set$oew, challenge_set$log_flown_distance)
# challenge_set$interaction_oew_log_flight_duration <- interaction(challenge_set$oew, challenge_set$log_flight_duration)

# Chatgpt suggestions
challenge_set$interaction_taxiout_time_flown_distance <- interaction(challenge_set$taxiout_time, challenge_set$flown_distance)
challenge_set$interaction_taxiout_time_great_circle_distance_adep_ades <- interaction(challenge_set$taxiout_time, challenge_set$great_circle_distance_adep_ades)
challenge_set$interaction_log_flown_distance_offblock_time_of_day <- interaction(challenge_set$log_flown_distance, challenge_set$offblock_time_of_day)
challenge_set$interaction_total_time_ascent_offblock_is_peak <- interaction(challenge_set$total_time_ascent, challenge_set$offblock_is_peak)
challenge_set$interaction_month_of_year_log_flown_distance <- interaction(challenge_set$month_of_year, challenge_set$log_flown_distance)
challenge_set$interaction_headwind_tailwind_bucket_altitude_range_bucket <- interaction(challenge_set$headwind_tailwind_bucket, challenge_set$altitude_range_bucket)
challenge_set$interaction_avg_humidity_bucket_total_time_ascent_bucket <- interaction(challenge_set$avg_humidity_bucket, challenge_set$total_time_ascent_bucket)
challenge_set$interaction_oew_flown_distance <- interaction(challenge_set$oew, challenge_set$flown_distance)
challenge_set$interaction_mtw_oew_diff_distance_bucket <- interaction(challenge_set$mtow_oew_diff, challenge_set$distance_bucket)
challenge_set$interaction_country_code_adep_elevation_ft_adep <- interaction(challenge_set$country_code_adep, challenge_set$elevation_ft_adep)
challenge_set$interaction_country_code_ades_elevation_ft_ades <- interaction(challenge_set$country_code_ades, challenge_set$elevation_ft_ades)
challenge_set$interaction_departure_traffic_distance_category <- interaction(challenge_set$departure_traffic, challenge_set$distance_category)
challenge_set$interaction_log_flown_distance_bucket_altitude_range_bucket <- interaction(challenge_set$log_flown_distance_bucket, challenge_set$altitude_range_bucket)
challenge_set$interaction_offblock_time_of_day_log_flown_distance_bucket <- interaction(challenge_set$offblock_time_of_day, challenge_set$log_flown_distance_bucket)
challenge_set$interaction_flight_duration_day_of_week <- interaction(challenge_set$flight_duration, challenge_set$day_of_week)
challenge_set$interaction_flight_duration_is_weekend <- interaction(challenge_set$flight_duration, challenge_set$is_weekend)
challenge_set$interaction_offblock_is_peak_month_of_year <- interaction(challenge_set$offblock_is_peak, challenge_set$month_of_year)
challenge_set$interaction_day_of_week_arrival_time_of_day <- interaction(challenge_set$day_of_week, challenge_set$arrival_time_of_day)
challenge_set$interaction_week_of_year_offblock_time_of_day <- interaction(challenge_set$week_of_year, challenge_set$offblock_time_of_day)
challenge_set$interaction_total_time_ascent_is_weekend <- interaction(challenge_set$total_time_ascent, challenge_set$is_weekend)
challenge_set$interaction_month_of_year_percentage_time_climb_4 <- interaction(challenge_set$month_of_year, challenge_set$percentage_time_climb_4)
challenge_set$interaction_avg_humidity_bucket_flight_duration <- interaction(challenge_set$avg_humidity_bucket, challenge_set$flight_duration)
challenge_set$interaction_departure_traffic_bucket_arrival_traffic_bucket <- interaction(challenge_set$departure_traffic_bucket, challenge_set$arrival_traffic_bucket)
challenge_set$interaction_flight_duration_predicted_tow <- challenge_set$flight_duration * challenge_set$lm_predicted_tow
challenge_set$interaction_flown_distance_predicted_tow <- challenge_set$flown_distance * challenge_set$lm_predicted_tow
challenge_set$interaction_flown_distance_r_squared <- challenge_set$flown_distance * challenge_set$lm_predicted_tow_r_squared
challenge_set$interaction_taxiout_time_r_squared <- challenge_set$taxiout_time * challenge_set$lm_predicted_tow_r_squared
challenge_set$interaction_aircraft_type_r_squared <- interaction(challenge_set$aircraft_type, challenge_set$lm_predicted_tow_r_squared)
challenge_set$interaction_r_squared_category_predicted_tow_category <- interaction(challenge_set$r_squared_category, challenge_set$predicted_tow_category)
```
## Designate Features

```{r}
# Define the list of columns to exclude
exclude_columns <- c("date", "callsign", "name_adep", "name_ades", "actual_offblock_time", "arrival_time", "tow", "flight_id", ".names", "median_track", "iqr_groundspeed", "median_u_component", "median_v_component", "median_wind_direction", "diff_track_vs_wind_direction", "wtc", "gps_code.x", "gps_code.y", "longitude_adep", "longitude_ades", "latitude_adep", "latitude_ades", "out_and_back_trip", "great_circle_distance")

# Get the names of all columns in the dataframe
all_columns <- names(challenge_set)

# Get the names of columns excluding the specified ones
features <- setdiff(all_columns, exclude_columns)

# View the selected features
print(features)
```

## Visualize Feature Distributions

```{r, fig.width=8, fig.height=25, out.width='100%'}
# # Load necessary libraries
# library(tidyr)  # For pivoting data
# 
# # Select numeric columns
# challenge_set %>% 
#   select(all_of(features)) %>%
#   select_if(is.numeric) %>%
# # Pivot longer to prepare for plotting
#   pivot_longer(everything(), names_to = "feature", values_to = "value") %>%
#   # Plot histograms for each numeric feature
#   ggplot(aes(x = value)) +
#     geom_histogram(bins = 30, fill = 'blue', color = 'white', alpha = 0.7) +
#     facet_wrap(~ feature, scales = "free", ncol = 3) +  # Creates one histogram per feature
#     theme_minimal() +
#     labs(title = "Histograms of Numeric Features", x = "Value", y = "Frequency")
```

```{r}
# Generate formula from vector
formula <- as.formula(paste("tow ~", paste(features, collapse = " + ")))
```

## Divide up Dataframe

```{r}
# Get only submissions set rows
submission_set <- challenge_set %>% filter(is.na(tow))
```

# Make Predictions

## Load Model

```{r}
# Load the saved model
rf_model <- readRDS("saved_models/random_forest_model_v18.rds")
```

## Make Prediction

```{r}
predictions <- predict(rf_model, data = submission_set)
```

## QA Prediction

```{r}
# submission_set %>% 
#   group_by(aircraft_type) %>% 
#   summarize(
#     count = n(),
#     avg_tow = mean(tow),
#     avg_dist = mean(flown_distance)
#   ) %>% 
#   arrange(desc(count))
```

```{r}
nrow(submission_set)
```

## Write Output

```{r}
submission_set$tow <- predictions$predictions
```

```{r}
team_name = 'team_amazing_forest'
version = 'v18'
team_id = 'c8bd88ad-040a-4521-91ca-3ee5f65443b0'

#results <- submission_set %>% arrange(original_order) %>% select(flight_id, tow)
results <- submission_set %>% select(flight_id, tow)

write_csv(results, paste0("estimates_final/", team_name, "_", version, "_", team_id, ".csv"))
```
